---
title: "Challenge 4: Millennials & Avocados"
author: "Janalyn Lao"
format: html
editor: visual
theme: cyborg
code-fold: show
---

# Introduction & Setup

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(here)

```

```{r}
#| label: import avocados
#| message: false

avocado <- read_csv(here::here("supporting_artifacts", 
                               "learning_targets", 
                               "challenge_4",
                               "avocado.csv"))
housing <- read_csv(here::here("supporting_artifacts", 
                               "learning_targets", 
                               "challenge_4",
                               "housing.csv"))

```

## Code Cleanup

### By Region

```{r}
#| label: code cleanup by region (needed for metro)
#| message: false

all_regions <- avocado |> 
  distinct(region)

avocado <- avocado |> filter(region != "TotalUS")

region <- data.frame(region = c("West", 
                                "Midwest", 
                                "SouthCentral", 
                                "Southeast",
                                "Plains",
                                "Northeast",
                                "Midsouth"))
avocado_region <- avocado |> 
  semi_join(region, 
            by = "region")

```

### By State

```{r}
#| label: code cleanup by state (needed for metro)
#| message: false

state <- data.frame(state = c("California",
                              "SouthCarolina",
                              "NewYork"))
avocado_state <- avocado |> 
  semi_join(state, 
            by = c("region" = "state"))


```

### By Metropolitan Area

```{r}
#| label: code cleanup by metro area
#| message: false

avocado_metro <- avocado |> 
  anti_join(region, 
            by = "region") |> 
  anti_join(state, 
            by = c("region" = "state"))

metro_check <- avocado_metro |> 
  distinct(region)

```

### Housing

```{r}
#| label: hosuing cleanup
#| message: false

housing <- housing |> 
  mutate(Year = as.character(Year))

```

# California Regions

```{r}
#| label: setup for CA region questions
#| message: false

cali <- data.frame(city = c("LosAngeles",
                            "SanDiego",
                            "Sacramento",
                            "SanFrancisco"))

avocado_ca <- avocado_metro |> 
  semi_join(cali, 
            by = c("region" = "city")) |> 
  separate(col = Date, 
           into = c('Year', 'Month', 'Day'), 
           sep = '-')

```

```{r}
#| label: combioning datasets and getting average prices
#| message: false

housing_avo <- avocado_ca |> 
  full_join(housing) |> 
  group_by(region, Year) |> 
  summarise(avg_avo = mean(AveragePrice),
            avg_housing = mean(`Average Housing Price`))

ggplot(data = housing_avo,
       mapping = aes(y = avg_avo,
                     x = avg_housing,
                     color = Year,
                     shape = region)) +
  geom_point(size = 3) +
  labs(x = "Housing Prices (thousands)", y = "Avocado Prices")

```

According to the scatter plot above, there is no data to suggest that Millennials can't afford houses because of avocado toast/the price of avocados. According to the joke, as housing prices increase, avocado sale prices should decrease so we should see a negative correlation of some kind, however, with how scattered this data looks, we can conclude that there is no such relation between the two prices.
