---
title: "Lab 4: Millennials & Avocados"
author: "Janalyn Lao"
format: 
  html:
    self-contained: true
editor: visual
theme: cyborg
code-fold: show
---

# Introduction & Setup

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(here)

```

1\) The avocado.csv dataset has the weekly retail scan information of Hass avocados from 2018. It includes information like the date of observation, region, total volume, average price, and total number of sales of specific PLU values. This data was taken from the Hass Avocado Board.

## Question 1

```{r}
#| label: import avocados
#| message: false

avocado <- read_csv(here::here("supporting_artifacts", "learning_targets", "avocado.csv"))

```

## Question 2

```{r}
#| label: code cleanup by region
#| message: false

all_regions <- avocado |> 
  distinct(region)

avocado <- avocado |> filter(region != "TotalUS")

region <- data.frame(region = c("West", 
                                "Midwest", 
                                "SouthCentral", 
                                "Southeast",
                                "Plains",
                                "Northeast",
                                "Midsouth"))
avocado_region <- avocado |> 
  semi_join(region, 
            by = "region")

```

## Question 3

```{r}
#| label: code cleanup by state
#| message: false

state <- data.frame(state = c("California",
                              "SouthCarolina",
                              "NewYork"))
avocado_state <- avocado |> 
  semi_join(state, 
            by = c("region" = "state"))


```

## Question 4

```{r}
#| label: code cleanup by metro area
#| message: false

avocado_metro <- avocado |> 
  anti_join(region, 
            by = "region") |> 
  anti_join(state, 
            by = c("region" = "state"))

metro_check <- avocado_metro |> 
  distinct(region)

```

## Question 5

```{r}
#| label: major region that sold the most organic, small Hass avocados in 2017
#| message: false

avocado_region |> 
  filter(type == "organic",
         year == 2017) |> 
  group_by(region) |> 
  summarize(tot_4046 = sum(`4046`)) |> 
  slice_max(order_by = tot_4046)

```

3\) The Northeast sold the most organic, small Hass avocados in 2017.

## Question 6

```{r}
#| label: month with highest volume avocado sales

avocado |> 
  separate(col = Date, 
           into = c('Year', 'Month', 'Day'), 
           sep = '-') |> 
  group_by(Month) |> 
  summarize(tot_sales = sum(`Total Volume`)) |> 
  slice_max(order_by = tot_sales)
  
```

4\) The highest volume of avocado sales was in February.

## Question 7

```{r}
#| label: metro areas with most total avocado sales

metro_sort <- avocado_metro |> 
  group_by(region) |> 
  summarise(tot_avo = sum(`Total Volume`)) |> 
  slice_max(order_by = tot_avo, n = 5)

metro_plot <- avocado |> 
  semi_join(metro_sort, by = "region")
  
ggplot(data = metro_plot,
         mapping = aes(x = region, y = `Total Volume`)) +
  geom_boxplot()

```

5\) The five metro area regions that sold the most total avocados are: GreatLakes, LosAngeles, DallasFtWorth, Houston, and PhoenixTucson.

## Question 8

```{r}
#| label: setup for CA region questions
#| message: false

cali <- data.frame(city = c("LosAngeles",
                            "SanDiego",
                            "Sacramento",
                            "SanFrancisco"))

avocado_ca <- avocado_metro |> 
  semi_join(cali, 
            by = c("region" = "city"))

```

## Question 9

```{r}
#| label: region with largest organic v. conventional diffrence
#| message: false

ggplot(data = avocado_ca,
       mapping = aes(x = region, 
                     y = AveragePrice,
                     color = type)) + 
  geom_boxplot() + 
  xlab("Average Price") + 
  ylab("California Region")

price_diff <- avocado_ca |> 
  group_by(region, type) |> 
  summarise(avg_price = mean(AveragePrice)) |> 
  pivot_wider(names_from = type, values_from = avg_price) |> 
  mutate(price_diff = abs(conventional - organic))
price_diff
  
```

6\) The price of organic vs. conventional avocados was the most different in San Francisco and Sacramento, with San Diego not far behind.

## Question 10

```{r}
#| label: recreating plot
#| warning: false
#| message: false

bag_plot <- avocado_ca |> 
  group_by(region, type) |> 
  summarise(across(`4046` : `4770`,
                   mean)) |> 
  pivot_longer(`4046` : `4770`, 
               names_to = "Avocado Size", 
               values_to = "Amount Sold") |> 
  mutate(`Avocado Size` = factor(`Avocado Size`,
                                 levels = c("4046", "4225", "4770"),
                                 labels = c("Small", "Large", "Extra Large")))

ggplot(data = bag_plot,
       mapping = aes(x = region, 
                     y = `Amount Sold`, 
                     col = `Avocado Size`, 
                     fill = `Avocado Size`)) +
  geom_col(position = "fill") +
  facet_wrap(~type) +
  xlab("Region of CA") + 
  ylab("Proportion of Mean Avocados Sold") + 
  theme(axis.text.x = element_text(vjust = grid::unit(c(-2, 0), "points")))
  
```
