---
title: "Challenge 7"
author: "Janalyn Lao"
format: 
  html:
    self-contained: true
editor: visual
theme: cyborg
code-fold: show
---

## Setup

```{r}
#| message: false

library(tidyverse)

```

```{r}
#| message: false

fish <- read_csv(here::here("supporting_artifacts", "learning_targets", "data", "BlackfootFish.csv"))

```

## Part 1

```{r}

out_of_bounds <- function(vec, min_val, max_val) {
  stopifnot(is.numeric(vec))
  stopifnot(is.numeric(min_val))    # check if min/max are numerics
  stopifnot(is.numeric(max_val))
  
  case_when(
    {{ vec }} < min_val ~ NA_real_,
    {{ vec }} > max_val ~ NA_real_,
    TRUE ~ {{ vec }}
  )
}

fish <- fish |> 
  mutate(length = out_of_bounds(length, 152, 508),
         weight = out_of_bounds(weight, 141, 1417)) |> 
  na.omit()



```

https://fieldguide.mt.gov/displaySpecies.aspx?family=Salmonidae

According to the Montana field guide, brown trout average 12-20in, rainbow trout average 12-17in, and westslope cutthroat trout average 6-16in, and bull trout average 6-12in. As a result, I want my minimum length to be 6in or about 152mm and maximum length to be 20in or 508mm.

https://www.dec.ny.gov/outdoor/9222.html

By using a ruler to get a general weight for my fish, the brown trout average should be 10-50oz, rainbow trout average should be 10-28oz, westslope cutthroat trout average should be 5-20oz, and bull trout average should be 5-10oz. As a result, I want my minimum weight to be 5oz or 141g and maximum weight to be 50oz or 1417g.

### Part 2

```{r}

condition_ind <- function(data, length, weight) {
  stopifnot(is.data.frame(data))
  
  data |> 
    mutate(condition_index = ({{ weight }} / {{ length }} ^ 3) * 100)
}

fish_CI <- condition_ind(fish, length, weight)

```

### Part 3

```{r}

avg_CI <- fish_CI |> 
  group_by(year, species, trip) |> 
  mutate(mean_CI = mean(condition_index))

  ggplot(data = avg_CI,
         mapping = aes(x = year, 
                       y = mean_CI)) +
    geom_line() +
    facet_grid(cols = vars(species), rows = vars(trip)) +
    labs(title = "Condition Index Of Each Trout Species By Trip",
         x = "Year Recorded",
         y = "")

# find the average CI across each species in each year

# https://www3.nd.edu/~steve/computing_with_data/13_Facets/facets.html

```
