---
title: "Challenge 3: Hip-Hop Lyrics"
author: "Janalyn Lao"
format: html
editor: visual
theme: cyborg
code-fold: show
---

## Setup

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(here)

```

```{r}
#| label: load data
#| message: false

hiphop <- read_csv(here::here("supporting_artifacts",
                              "continued_learning",
                              "revisions",
                              "ch3_revision",
                              "hiphop.csv"))

```

## Published Comparisons

```{r}
#| label: create new dataset for male vs female
#| message: false

hiphop |> 
  group_by(sex) |> 
  summarise(
    across(intl : unclassifiable, mean)
  ) |> 
  summarise(
    across(intl : unclassifiable, diff)
  )

```

Between males and females, the highest difference in mean reported artists was in rock.

```{r}
#| label: create new dataset for white vs non-white
#| message: false

hiphop |> 
  mutate(ethnic = if_else(ethnic != "white", "non-white", "white")) |> 
  group_by(ethnic) |> 
  summarise(
    across(intl : unclassifiable, mean)
  ) |> 
  summarise(
    across(intl : unclassifiable, diff)
  )

```

Between white and non-white subjects, the highest difference in mean reported artists was in international music.

[Great work! You are on the right track! The next step is to find the differences in these means. Could you use another `summarize()` step to get these differences?]{style="color:red"}

[Where is the code output to confirm your summary?]{style="color:red"}

## Concerns of the Study

One concern I have is for the first move question, the name and ethnicity of the persons involved were asked. I wonder if a person listed someone with a stereotypical African American name, it would somehow affect the data. For example, I believe I know a fair amount of AAE terms because of the community I grew up in, however, I would list names like "Phoebe, Asian" or "Dennis, Asian" for this question, which wouldn't really do anything to enforce any data.

Another concern I have is with the second move question, it seems like it is implied that being friends/acquaintances with an African American person would somehow lead to more insight on a person's data outcome. There are many African American individuals who don't listen to hip-hop or speak in the AAE dialect so this question seems to generalize them all together.
